Wigglyhedra

### 2024-02-28 draft/wigglyhedra.tex
minor changes
fix Definition 31 + remark universal associahedra

@@

\begin{definition}
\label{def:bijection2}
For a wiggly permutation~$\sigma$ of~$[2n]$, we define~$\Psi(\sigma) \eqdef \set{\alpha(\sigma, k)}{k \in [2n-1]}$, where for~$k \in [2n-1]$, we have~$\alpha(\sigma, k) \eqdef \big( i, j, {]i,j[} \cap \sigma([k]), {]i,j[} \ssm \sigma([k]) \big)$ with
\[
i \eqdef \max \{0\} \cup \set{i \in [n]}{2i-1 \in \sigma([k]) \not\ni 2i} 
\;\text{and}\;
j \eqdef \min \{n+1\} \cup \set{j \in [n]}{2j \in \sigma([k]) \not\ni 2j-1} \!.
\]
\end{definition}

@@

Observe now that~${\b{p}(T') - \b{p}(T)}$ is orthogonal to the $(2n-2)$-dimensional cone of the wiggly fan generated by the $\b{g}$-vectors of the wiggly arcs in~$T \cap T'$, and points from the side containing~$\b{g}(\alpha)$ to the side containing~$\b{g}(\alpha')$.
Hence, $\b{p}(T') - \b{p}(T)$ is a positive multiple of~$-\b{c}(\alpha, T)$.
From \cref{def:cvectors}, we observe that
\begin{itemize}
\item $\dotprod{\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} = 1/2$ if~$i = \bar u \ne \bar v$, $-1/2$ if~$i = \bar v \ne \bar u$, and $0$ otherwise,
\item $\dotprod{\b{c}(\alpha, T)}{\b{e}_j}| \le 1/2$ for all~$j \in [2n]$.
\end{itemize}
We now distinguish two situations.
If~$\bar u \ne \bar v$, then
\[
\dotprod{-\b{c}(\alpha, T)}{\b{\omega}} = \sum_{i \in [n]} 4ni \dotprod{-\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} + \sum_{j \in [2n]} j \dotprod{-\b{c}(\alpha, T)}{\b{e}_j} \ge 2n(\bar v - \bar u) + n > 0.
\]
If~$\bar u = \bar v$, then~$u = 2k-1$ and~$v = 2k$ for some~$k \in [n]$, and we have
\[
\dotprod{\b{c}(\alpha, T)}{\b{\omega}} = \sum_{i \in [n]} 4ni \dotprod{-\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} + \sum_{j \in [2n]} j \dotprod{-\b{c}(\alpha, T)}{\b{e}_j} = 1/2 > 0.
\]
We conclude in both cases that~$\dotprod{\b{c}(\alpha, T)}{\b{\omega}} > 0$, so that we recover the increasing flip graph by orienting the graph of the wigglyhedron~$\wigglyhedron_n$ in direction~$\b{\omega}$.
\end{proof}

@@

>>>>

Analogies to cluster algebras and gentle algebras.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{definition}
\label{def:bijection2}
For a wiggly permutation~$\sigma$ of~$[2n]$, we define~$\Psi(\sigma) \eqdef \set{\alpha(\sigma, k)}{k \in [2n-1]}$, where for~$k \in [2n-1]$, we have~$\alpha(\sigma, k) \eqdef ( i, j, A, B)$ with
\begin{align*}
i & \eqdef \max \{0\} \cup \set{i \in [n]}{2i-1 \in \sigma([k]) \not\ni 2i}, \\
j & \eqdef \min \{n+1\} \cup \set{j \in [n]}{2j \in \sigma([k]) \not\ni 2j-1}, \\
A & \eqdef \set{\ell \in {]i,j[}}{\{2\ell-1, 2\ell\} \subseteq \sigma([k])}, \\
B & \eqdef \set{\ell \in {]i,j[}}{\{2\ell-1, 2\ell\} \cap \sigma([k]) = \varnothing}.
\end{align*}
\end{definition}

@@

Observe now that~${\b{p}(T') - \b{p}(T)}$ is orthogonal to the $(2n-2)$-dimensional cone of the wiggly fan generated by the $\b{g}$-vectors of the wiggly arcs in~$T \cap T'$, and points from the side containing~$\b{g}(\alpha)$ to the side containing~$\b{g}(\alpha')$.
Hence, $\b{p}(T') - \b{p}(T)$ is a positive multiple of~$-\b{c}(\alpha, T)$.
From \cref{def:cvectors}, we observe that
\begin{itemize}
\item $\dotprod{\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} = 1/2$ if~$i = \bar u \ne \bar v$, $-1/2$ if~$i = \bar v \ne \bar u$, and $0$ otherwise,
\item $|\dotprod{\b{c}(\alpha, T)}{\b{e}_j}| \le 1/2$ for all~$j \in [2n]$.
\end{itemize}
We now distinguish two situations.
If~$\bar u \ne \bar v$, then
\[
\dotprod{-\b{c}(\alpha, T)}{\b{\omega}} = \sum_{i \in [n]} 4ni \dotprod{-\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} + \sum_{j \in [2n]} j \dotprod{-\b{c}(\alpha, T)}{\b{e}_j} \ge 2n(\bar v - \bar u) - n > 0.
\]
If~$\bar u = \bar v$, then~$u = 2k-1$ and~$v = 2k$ for some~$k \in [n]$, and we have
\[
\dotprod{\b{c}(\alpha, T)}{\b{\omega}} = \!\! \sum_{i \in [n]} 4ni \dotprod{-\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} + \!\! \sum_{j \in [2n]} j \dotprod{-\b{c}(\alpha, T)}{\b{e}_j} = \frac{2k}{2}-\frac{2k-1}{2} = 1/2 > 0.
\]
We conclude in both cases that~$\dotprod{\b{c}(\alpha, T)}{\b{\omega}} > 0$, so that we recover the increasing flip graph by orienting the graph of the wigglyhedron~$\wigglyhedron_n$ in direction~$\b{\omega}$.
\end{proof}

@@

\begin{remark}
In their work on polytopal realizations of $\b{g}$-vector fans of finite type cluster algebras~\cite{HohlwegPilaudStella}, C.~Hohlweg, V.~Pilaud and S.~Stella already constructed a \defn{universal associahedron}~$\polytope{U}_n$ which somehow contains all Cambrian fans (actually all $\b{g}$-vector fans from any initial cluster seed).
More precisely, the $\delta$-Cambrian fan for each~$\delta \in \{+,-\}^n$  appears as a section of the normal fan of the universal associahedron~$\polytope{U}_n$, which is a $(n^2+n-4)/2$-dimensional polytope.
In contrast, \cref{prop:CambrianFan} states that the $\delta$-Cambrian fan for each~$\delta \in \{+,-\}^n$ appears as a link of the normal fan of the wigglyhedron~$\wigglyhedron_n$, which is a $(2n-1)$-dimensional polytope.
There seem to be no connection between these two universal polytopes.
However, this certainly raises the question to construct analogues of the wigglyhedron for all finite Coxeter groups.
\end{remark}
